---
output: html_document
runtime: shiny
---

---
title: "PSY6422 Project"
author: "E Irving"
date: "13/04/2022"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) ##packages needed
library(here)
library(tidyverse)
library(shiny)
```

# Data Origins 

### Where is the data from?
The Office for National Statistics (ONS)

Citation: Office for National Statistics. (2022,April). Personal Well-being estimates by local authority. https://www.ons.gov.uk/datasets/wellbeing-local-authority/editions/time-series/versions/2


### How was it collected?
This data was collected through the Annual Population Survey (APS)

### What do the variables mean?
The variables in the dataset include: 'Anxiety', 'Life Satisfaction', 'Worthwhile', and 'Happiness'. These variables are the four factors used to measure personal well-being. 
For 'Anxiety', higher scores indicate lower personal well-being. 
For 'Life satisfaction', 'Worthwhile', and 'Happiness', high scores equal high personal well-being. 

#### Show first few rows of raw data if possible

```{r data import}
pwbestlocalauth <- read.csv(here("data", "pwbest_localauth_filtered.csv"))
#head(pwbestlocalauth) 
```

```{r showing first few rows of raw data}
head(pwbestlocalauth, 4)
```

# Research Questions
What question(s) does my visualisation attempt to address?
 
1) How has personal well-being across UK countries, and the UK overall, fluctuated over recent years?

2) On average, Which UK countries appear to have a higher level of personal well-being each year? 

# Data Preparation

```{r datawrangling}
#rename: change column/variable names to something appropriate
pwbestlocalauth <- pwbestlocalauth %>% 
  rename(Geography = Title, pwb = X, '2016-17' = X.2, '2017-18' = X.6, '2018-19' = X.10, '2019-20' = X.14, '2020-21' = X.18)

# use select to only look at some columns
pwbestlocalauth <- pwbestlocalauth %>% 
  select(Geography, pwb, '2016-17', '2017-18','2018-19', '2019-20', '2020-21')
#head(pwbestlocalauth)

#remove rows from dataset
pwbestlocalauth <- pwbestlocalauth[-c(1,2),]
head(pwbestlocalauth, 3)
```

## Summary statistics - grouped variables etc. 

```{r group statistics}


```


# Visualisations
-inlude a document explaining any motivation (although good graph labelling is better than an explanation in the accompanying text.)
The code for producing these graphs. 

## Visualisation 1

A scatter plot to show the personal well-being data for the year 2016-17:

```{r scatterplot}

cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7") #use of colourblind palette 

#quick visualization test of datapoints
#p <- ggplot(pwbestlocalauth,aes(x=pwb,y=`2016-17`,color = Geography))
#p + geom_point(size=2.5) + geom_smooth(alpha=0.3, method = 'lm') + labs(x = "Personal Wellbeing Measures",y = "Scores", title = "Personal Wellbeing Scores in 2016-2017") + #scale_colour_manual(values=cbPalette)
#ggsave(here("figs", "16-17plot.png"))
```

## Visualisation 2

Shiny interactive graph:

```{r Shiny}

ui <- shinyUI(fluidPage(
  titlePanel("Personal Wellbeing Scores across the UK"),
  sidebarLayout(
    sidebarPanel(
      radioButtons(inputId = "year", #created a button panel for the user to select the year they wish to view
                   label = "Please select a year:",
                   choices=c("2016-17" = "2016-17", "2017-18" = "2017-18","2018-19" = "2018-19", "2019-20" = "2019-20","2020-21" = "2020-21"),
                   
                   selected = "2016-17" #automatically set on the year 2016-2017
      ),
      radioButtons(inputId="filetype", label="Select the file type to download", choices = list("png","pdf")),#buttonpanel to choose file type to download plot
      downloadButton(outputId = "down", label = "Download plot")),
    
    mainPanel(
      plotOutput("scatterplot"),
      textOutput("text"),
      textOutput("text2"),
      textOutput("text3")
    )
)
))


#instructions for server of what to display
server <- shinyServer(function(input,output){
  output$scatterplot <- renderPlot({#ggplot used to create a scatterplot of the data depending on year picked by user
   ggplot(pwbestlocalauth, aes(x= pwb, y=get(input$year), color = Geography)) + geom_point(size=3, position=position_jitter(h=.15,w=.12)) +  #jitter was used as points overlapped
      geom_smooth(alpha=0.3, method = 'lm') + 
      #customisation of plot for better visualisation
      labs(x = "Personal Wellbeing Measures", y = "Scores", title = "Personal Wellbeing Scores") + scale_colour_manual(values=cbPalette) + theme_light() + theme(text=element_text(color="black", size=11), legend.position = c("top"), legend.direction = "horizontal", axis.title.x=element_text(vjust=-1.5),
                              legend.justification = 0.5, legend.text=element_text(size=11,color="black"), axis.line=element_line(color="black", size=0.2), axis.line.y=element_blank(),panel.grid.major=element_line(color="gray50", size=0.5), panel.grid.major.x = element_blank())
  })
  
  #text used to display a key for the data
  output$text <- renderText({"Personal WellBeing (PWB) Scores Key:"})
  output$text2 <- renderText({"Low Anxiety=High PWB"})
  output$text3 <- renderText({"Low Happiness,Low Satisfaction, Low Worthwhile = Low PWB"})
  
  output$down <- downloadHandler( #code to enable download of specific plot
    filename = function(){
     paste("pwb",input$year,input$filetype,sep='.') #code to name file being downloaded 
    },
    content = function(file){
      if(input$filetype == "png")
        png(file)
      else
        pdf(file) 
      plot(ggplot(pwbestlocalauth, aes(x= pwb, y=get(input$year), color = Geography)) + geom_point(size=3, position=position_jitter(h=.15,w=.12)) +  geom_smooth(alpha=0.3, method = 'lm') + labs(x = "Personal Wellbeing Measures", y = "Scores", title = "Personal Wellbeing Scores") + scale_colour_manual(values=cbPalette) + theme_light() + theme(text=element_text(color="black", size=11), legend.position = c("top"), legend.direction = "horizontal", axis.title.x=element_text(vjust=-1.5),
                              legend.justification = 0.5, legend.text=element_text(size=11,color="black"), axis.line=element_line(color="black", size=0.2), axis.line.y=element_blank(),panel.grid.major=element_line(color="gray50", size=0.5), panel.grid.major.x = element_blank()))
  dev.off()
    }
  )
  
  

})

shinyApp(ui, server)
```

Interpretation of Visualisation:

Based on the Shiny graphs, it is evident that overall, Northern Ireland has higher PWB scores across all measures, in comparison to other UK countries. In comparison, Wales seems to have the worst PWB scores across all measures. 


# Summary

What do I think the graph means?

What have I learnt whilst doing this project?

brief thoughts on what you have learnt, what you might do next if you had more time/more data
This practice project has enabled me to produce simple ggplots with my data, as well as creating a shiny interface where different graphs can be shown based on the year a user is interested in.



